\section{Dynamic Programming}
% ============================================================

\begin{desc}
\textbf{Dynamic Programming} solves problems by breaking them into overlapping
subproblems and storing results to avoid recomputation.\\
\textbf{Two key properties}:
(1) \textbf{Optimal substructure}---optimal solution contains optimal solutions to subproblems.
(2) \textbf{Overlapping subproblems}---same subproblems recur many times.\\
\textbf{Approach}: identify state $\to$ write recurrence $\to$ determine evaluation order $\to$
code bottom-up (or top-down with memoization).
\end{desc}

\subsection{0/1 Knapsack}
\begin{desc}
Given $n$ items with weights $w_i$ and values $v_i$, and a knapsack of capacity $C$,
maximize total value without exceeding capacity. Each item is either \textbf{taken or not} (no fractions).\\
State: $\text{dp}[i][j]$ = max value using first $i$ items with capacity $j$.\\
Recurrence: $\text{dp}[i][j] = \max(\text{dp}[i-1][j],\; \text{dp}[i-1][j-w_i] + v_i)$ if $w_i \le j$.
\end{desc}
\complexity{$O(nC)$}{$O(nC)$, or $O(C)$ with space optimization}
\begin{lstlisting}
static int knapsack(int[] w, int[] v,
    int cap) {
  int n = w.length;
  var dp = new int[n + 1][cap + 1];
  for (int i = 1; i <= n; i++)
    for (int j = 0; j <= cap; j++) {
      dp[i][j] = dp[i-1][j]; // skip item i
      if (w[i-1] <= j)       // take item i
        dp[i][j] = Math.max(dp[i][j],
          dp[i-1][j - w[i-1]] + v[i-1]);
    }
  return dp[n][cap];
}
\end{lstlisting}

\subsection{Longest Common Subsequence (LCS)}
\begin{desc}
Find the longest subsequence common to two strings. A \textbf{subsequence} keeps
order but can skip characters (not necessarily contiguous).\\
State: $\text{dp}[i][j]$ = LCS length of $a[0..i{-}1]$ and $b[0..j{-}1]$.\\
Recurrence: If $a[i{-}1] = b[j{-}1]$: extend by 1. Otherwise: take max of skipping
from either string.\\
Useful for text comparison, diff algorithms, edit distance variations.
\end{desc}
\complexity{$O(mn)$}{$O(mn)$}
\begin{lstlisting}
static int lcs(String a, String b) {
  int m = a.length(), n = b.length();
  var dp = new int[m + 1][n + 1];
  for (int i = 1; i <= m; i++)
    for (int j = 1; j <= n; j++)
      dp[i][j] =
        a.charAt(i-1) == b.charAt(j-1)
          ? dp[i-1][j-1] + 1
          : Math.max(dp[i-1][j],
                     dp[i][j-1]);
  return dp[m][n];
}
\end{lstlisting}

\subsection{Coin Change (Unbounded Knapsack)}
\begin{desc}
Find the \textbf{minimum number of coins} to make a target amount, with unlimited
supply of each denomination.\\
State: $\text{dp}[j]$ = min coins to make amount $j$.\\
Recurrence: For each coin $c$: $\text{dp}[j] = \min(\text{dp}[j],\; \text{dp}[j-c] + 1)$.\\
Can be adapted to count ways, find all combinations, etc.
\end{desc}
\complexity{$O(n \cdot \text{amt})$}{$O(\text{amt})$}
\begin{lstlisting}
static int coinChange(int[] coins, int amt) {
  var dp = new int[amt + 1];
  Arrays.fill(dp, amt + 1); // "infinity"
  dp[0] = 0;
  for (int c : coins)
    for (int j = c; j <= amt; j++)
      dp[j] = Math.min(dp[j], dp[j-c] + 1);
  return dp[amt] > amt ? -1 : dp[amt];
}
\end{lstlisting}

\subsection{Longest Increasing Subsequence (LIS)}
\begin{desc}
Find the length of the longest \textbf{strictly increasing} subsequence.\\
The $O(n \log n)$ approach maintains an array \texttt{tails} where \texttt{tails[i]}
is the smallest tail element of all increasing subsequences of length $i+1$.
For each element, binary search its position and update.\\
Useful for sequence optimization and patience sorting.
\end{desc}
\complexity{$O(n \log n)$}{$O(n)$}
\begin{lstlisting}
static int lis(int[] a) {
  var tails = new ArrayList<Integer>();
  for (int x : a) {
    int pos = Collections
      .binarySearch(tails, x);
    if (pos < 0) pos = -(pos + 1);
    if (pos == tails.size()) tails.add(x);
    else tails.set(pos, x);
  }
  return tails.size();
}
\end{lstlisting}

\subsection{Subset Sum (Exact Match)}
\begin{desc}
Determine if any subset of items sums to exactly a target value.
This is the decision version of 0/1 knapsack (exists vs. optimize).\\
State: $\text{dp}[j]$ = true if sum $j$ is achievable.\\
Common in partition problems, exact matching, and backpack variants.
\end{desc}
\complexity{$O(n \cdot \text{target})$}{$O(\text{target})$}
\begin{lstlisting}
static boolean subsetSum(int[] a, int target){
  var dp = new boolean[target + 1];
  dp[0] = true;
  for (int x : a)
    // Iterate backwards to avoid using same item twice
    for (int j = target; j >= x; j--)
      if (dp[j - x]) dp[j] = true;
  return dp[target];
}
// To reconstruct which items: track parent pointers
// or use 2D dp and backtrack
\end{lstlisting}

\subsection{Bitmask DP}
\begin{desc}
When the number of elements is small, represent the ``used'' set as a bitmask.
State: $\text{dp}[\text{mask}]$ = best result using the subset encoded by mask.\\
\textbf{Applications}: assignment problems, traveling salesman (TSP), set cover,
Hamiltonian path, puzzle states.\\
For $n \le 20$: use \texttt{int} or \texttt{long} bitmask (fits in memory: $2^{20} \approx 1$M states).\\
For $n \le 64$: use \texttt{long} bitmask.\\
For $n > 64$: use \texttt{BitSet} (see Java API section).
\end{desc}
\begin{lstlisting}
// Example: assign n tasks to n workers (n <= 20)
// dp[mask] = min cost when tasks in "mask" done
var dp = new int[1 << n];
Arrays.fill(dp, Integer.MAX_VALUE);
dp[0] = 0;

for (int mask = 0; mask < (1<<n); mask++) {
  if (dp[mask] == Integer.MAX_VALUE) continue;
  int worker = Integer.bitCount(mask);
  for (int task = 0; task < n; task++) {
    if ((mask & (1<<task)) != 0) continue;
    int next = mask | (1 << task);
    dp[next] = Math.min(dp[next],
      dp[mask] + cost[worker][task]);
  }
}
// answer = dp[(1<<n) - 1]

// For n > 64, use BitSet instead:
var dp2 = new HashMap<BitSet, Integer>();
var initial = new BitSet();
dp2.put(initial, 0);
// ... iterate through states using BitSet operations
\end{lstlisting}

\subsection{Grid DP}
\begin{desc}
Count paths or find optimal value traversing a grid from top-left to bottom-right
(moving only right/down, or with other directional constraints).\\
State: $\text{dp}[i][j]$ = best value or count to reach cell $(i,j)$.\\
Common in grid optimization, robot path counting, matrix chain problems.
\end{desc}
\begin{lstlisting}
// Max sum path (right/down only)
var dp = new long[R][C];
dp[0][0] = grid[0][0];

// Fill first row and column
for (int i = 1; i < R; i++)
  dp[i][0] = dp[i-1][0] + grid[i][0];
for (int j = 1; j < C; j++)
  dp[0][j] = dp[0][j-1] + grid[0][j];

// Fill rest
for (int i = 1; i < R; i++)
  for (int j = 1; j < C; j++)
    dp[i][j] = Math.max(dp[i-1][j],
      dp[i][j-1]) + grid[i][j];

// answer = dp[R-1][C-1]
\end{lstlisting}

% ============================================================
